---
title: "New York State Temperature and Precipitation, 1981 - 2010"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
---

```{r setup, include=FALSE}
library(flexdashboard)
library(tidyverse)
library(viridis)
library(p8105.datasets)
library(plotly)
library(rnoaa)
library(zoo)
library(lubridate)
library(patchwork)
```

The highest temperature ever recorded in New York State is 43 degrees Celsius and the lowest recorded is -47 degrees Celsius, according to the National Climatic Data Center. Temperature observations above and below -47 and 43 degrees Celsius were removed, therefore, under the assumption that values outside that range reflect inaccurate measurements. Precipitation observations that were greater than the highest daily precipitation seen in New York state (~14 inches) were also removed.  

```{r data cleaning, echo = FALSE, include = FALSE}
#NOAA precip data
 data("ny_noaa")

noaa_data_cleaned <- ny_noaa %>% 
  separate(date, c("y", "m", "d")) %>% 
  mutate_at(vars(y:tmin), funs(as.numeric)) %>% 
  mutate_at(vars(prcp, tmax:tmin), funs(./10)) %>% #convert precip and temp data into mm and deg C, respectively
  rename(year = y) %>% 
  select(-snwd) %>% 
  filter(tmax < 43 & tmax > -47) %>% #remove inaccurate max temp observations 
  filter(tmin > -47 & tmin < 43) %>% #remove inaccurate min temp observations
  filter(prcp < 280) #remove inaccurate precip observations 
```


Column {data-width=650}
-----------------------------------------------------------------------

### Chart A

```{r}
temp_data <- noaa_data_cleaned %>% 
  group_by(year,m, d) %>% 
  summarise(avg_high = mean(tmax, na.rm = TRUE), avg_low = mean(tmin, na.rm = TRUE)) %>% 
  na.omit() %>% 
  mutate(time = m + (d/31))
  
temp_plot_high <-  plot_ly(temp_data, x = ~time, y = ~avg_high, type = "scatter", mode = "lines+markers", 
                     alpha = 0.8, 
                     color = ~year) %>% 
                layout(
                  title = "Average Daily Maximum Temperature in New York State, 1981 - 2010", 
                  xaxis = list(title = "Day"), 
                  yaxis = list(title = "Average Max Temperature (deg. C)")
                  )

temp_plot_low <- plot_ly(temp_data, x = ~time, y = ~avg_low, type = "scatter", mode = "lines+markers", 
                     alpha = 0.4, 
                     color = ~year) %>% 
                layout(
                  title = "Average Daily Minimum Temperature in New York State, 1981 - 2010", 
                  xaxis = list(title = "Day"), 
                  yaxis = list(title = "Average Min Temperature (deg. C)")
                  )

temp_plot_high
temp_plot_low
```

Column {data-width=350}
-----------------------------------------------------------------------

### Chart B

```{r}
precip_data <- noaa_data_cleaned %>% 
  group_by(year,m) %>% 
  summarise(total_precip = sum(prcp, na.rm = TRUE), mean_precip = mean(prcp, na.rm = TRUE)) %>% 
  na.omit() %>% 
  mutate(month = as.factor(m)) 

precip_plot <- plot_ly(precip_data, y = ~mean_precip, color = ~month, type = "box", colors = "Set2") %>% 
  layout(title = "Average Monthly Precipitation in New York State, 1981 -2010", 
         xaxis = list(title = "Month"),
         yaxis = list(title = "Precipitation (mm)"))

precip_plot

```

### Chart C

```{r snow}
snow_data <- noaa_data_cleaned %>% 
  group_by(year, m, d) %>% 
  summarise(total_snow = sum(snow, na.rm = TRUE)) %>% 
  mutate(snow_days = ifelse(total_snow > 0, 1, 0)) %>% 
  group_by(year) %>% 
  summarise(num_days_snow = sum(snow_days, na.rm = TRUE)) 

snow_plot <- plot_ly(snow_data, x = ~year, y = ~num_days_snow, color = ~year, type = "bar") %>% 
  layout(title = "Number of Days with Snow in New York State, 1981 - 2010", 
         xaxis = list(title = "Year"), 
         yaxis = list(title = "Number of Days"))

snow_plot
```

```{r}
rmarkdown::render("noaa_plots.Rmd", output_format = "flex_dashboard")
```

